Parsing Postfix log files

### 2007-12-04 doc/logparser.tex
Improved the first half of the review of other parsers; it's more compare and contrast than review now (or getting there, at least).

% Useful stuff for math mode.
\usepackage{amstext}
% Wrapping of URLs.  This is used in the bibliography instead of \url{}.
\usepackage{url}
\newcommand{\parsername}{ASO}
% Include images
\usepackage[final]{graphicx}
% Change how nested enumerate environments are labelled.
\renewcommand{\labelenumii}{\roman{enumii}:}
\renewcommand{\refname}{Bibliography}
% Add the bibliography into the table of contents.
\usepackage[section,numbib]{tocbibind}

@@

        Pflogsumm produces a report designed for troubleshooting, and does
        not support saving any data; parsing concentrates on extracting
        only the data required to produce the report.  Both the parsing and
        reporting are difficult to extend, as it is a specialised tool.  It
        does not attempt to correlate log lines by queueid or \pid{}, nor
        does it need to deal with the complications encountered during this
        project.  Pflogsumm produces a useful report, and successfully
        dealt with all the Postfix logs tested with.

@@

        Sawmill is a general purpose commercial product which parses 687
        log file formats (correct as of 2007/04/15) and produces reports.
        Its data extraction facilities are quite limited, though it does
        extract three different sets of data for Postfix (one is beta as of
        2007/04/15), but they do not appear to be interlinked, nor does it
        save sufficient data for the purposes of this project.  No attempt
        is made to correlate log lines or deal with the difficulties
        documented in sections~\refwithpage{complications}
        and~\refwithpage{additional complications}.\footnote{If any attempt
        is made there is no reference to it in the documentation available
        on the website.} The source code is available in an obfuscated form
        only (presumably for a fee), and the product is quite expensive, as
        it requires a license per report which is to be generated.  The web
        interface allows creation of dynamic reports based on any field,
        but due to the associated cost the author has not experimented with
        it.

@@

        (Last checked 2007/09/26.)

@@

        Splunk aims to index all an organisation's logs, providing a
        centralised view capable of searching and correlating diverse log
        sources.  The web interface allows complicated searches, providing
        statistics and graphs in real time.  Saved searches can be
        configured to run periodically and the results emailed to a
        recipient or sent to a shell script, which presumably can publish
        the results as required (though possibly without the graphs and
        detailed statistics).  In the specific case of parsing Postfix
        logs, Splunk extracts some standard fields: to and from addresses,
        HELO hostname, date, host the logs were collected from, and
        protocol.  It parses the standard syslog fields at the beginning of
        the line, and extracts any \texttt{key=<value>} pairs; these are
        the fields listed above.  Searches can be based on the extracted
        fields, and all text in the line is also available for searching.
        Many types of reports are available, though all are variations of a
        bar or pie chart, with the exception of bubble and heatmap graphs.
        It is very easy to drill down through the graphs to extract a
        portion of the data (e.g.\ select the hour with the largest number
        of events, then select a particular host, and finally a specific
        address).  It is not possible to search on partial words, though
        the web interface does provide and refine a list of possible search
        terms as you type.  The searches are quite fast, though as the free
        version has a limit on the amount of data indexed per day the
        volume of data to be searched was merely 45mb; the cheapest
        licensed version costs \$6000, and it still limits the volume of
        data indexed per day.

        The interface is optimised for interactive rather than automated
        queries and it does not appear to be possible to write independent
        tools to utilise the Splunk database.  Splunk is a generic tool, so
        it lacks any Postfix specific support over and above extracting the
        \texttt{key=<value>} fields; most importantly it makes no attempt
        to correlate log lines by queueid or \pid{}, nor to handle any of
        the myriad complications discussed in this document.  Correlation
        of log lines is left up to the user, and Splunk certainly makes the
        job easier than using grep, awk and similar Unix tools unless the
        user has a very high proficiency with those tools.

@@

        Isoqlog produces a report listing the number of mails sent by each
        unique sender address, and separately the total number of bytes
        transferred; both reports are produced for daily, monthly and
        annual time spans, but only for the domains listed in its
        configuration file.  It appears to ignore all log lines except for
        those for the current day, though it does maintain a record of data
        previously extracted, which the newly extracted data is merged into
        (no information is provided on the format of the data store).  It
        doesn't utilise rejection log lines in any way, so is unsuitable
        for the purposes of this project.  It doesn't handle any of the
        complications discussed in this document, doesn't gather the
        breadth of data required for this project, and ignores the majority
        of log lines produced by Postfix.

@@

        AWStats will produce simple graphs for many different services, but
        supporting many different services without special purpose code
        restricts it to supporting the \LCD{}.  The data it will extract
        from an \MTA{} log file is limited to: \newline \tab{} time2,
        email, email\_r, host, host\_r, method, url, code and bytesd.
        \newline  There does not seem to be an explanation of those fields
        in the documentation.  AWStats has no special purpose code to deal
        with the intricacies of Postfix logs, in fact it operates by
        coercing Postfix logs into Apache\footnote{The Apache web server is
        the most popular HTTP server in use for the past 10 years; more
        information is available at \url{http://httpd.apache.org/}} format
        log files, for analysis by AWStats' HTTP log file parser.  The
        converting parser only deals with a small portion of the log lines
        generated by Postfix, and though it does correlate log lines by
        queueid, it does not deal with any of the other complications
        described in this document.

@@

>>>>

% Useful stuff for math mode.
\usepackage{amstext}
% Wrapping of URLs.  This is used in the bibliography instead of \url{}.
\usepackage{url}
\newcommand{\parsername}{ASO}
\newcommand{\parsernames}{ASO's}
% Include images
\usepackage[final]{graphicx}
% Change how nested enumerate environments are labelled.
\renewcommand{\labelenumii}{\roman{enumii}:}
\renewcommand{\refname}{Bibliography}
% Add the bibliography into the table of contents.
\usepackage[section,numbib]{tocbibind}

@@

        Pflogsumm produces a report designed for troubleshooting, rather
        than for in-depth analysis.  It does not support saving any data,
        nor does it extract more data than is required for producing the
        report.  Both the parsing and reporting are difficult to extend as
        it is a specialised tool, unlike the easily extensible design of
        \parsername{}.  It does not attempt to correlate log lines by
        queueid or \pid{}, nor does it need to deal with the complications
        encountered during this project.  Pflogsumm produces a useful
        report, and successfully dealt with all the Postfix logs tested
        with, but is not a suitable base for this project.

@@

        Sawmill is a general purpose commercial product which parses 687
        log file formats (correct as of 2007/12/04) and produces reports.
        Its data extraction facilities are quite limited, though it does
        extract three different sets of data for Postfix (one is beta as of
        2007/12/04), but they do not appear to be interlinked, nor does it
        save sufficient data for the purposes of this project.  No attempt
        is made to correlate log lines or deal with the difficulties
        documented in sections~\refwithpage{complications}
        and~\refwithpage{additional complications}.\footnote{If any attempt
        is made there is no reference to it in the documentation available
        on the website.} The source code is available in an obfuscated form
        only (presumably for a fee), and the product is quite expensive, as
        it requires a license per report which is to be generated; in
        contrast \parsername{} is free and the code is freely available.
        The web interface allows creation of dynamic reports based on any
        field, but due to the associated cost the author has not
        experimented with it.  Presumably as the company charges per report
        the user wishes to generate the data store (if there is one) is
        deliberately inaccessible and undocumented to prevent the user
        bypassing the program and generating their own reports.

@@

        Even if the three data sets were linked together Sawmill would
        extract less data than \parsername{}, and it does not appear to
        extract data about rejections except when the rejection is caused
        by an \RBL{} check.

        (Last checked 2007/12/04.)

@@

        Splunk aims to index all an organisation's logs, providing a
        centralised view capable of searching and correlating diverse log
        sources.  The web interface supports complicated searches,
        providing statistics and graphs in real time, a facility not
        provided by \parsername{} (report generation has been deferred to a
        subsequent program).  Saved searches\footnote{The author was unable
        to save searches, though that may have been due to limitations in
        the free version.} can be run periodically and the results emailed
        to a recipient or sent to a shell script, which presumably can
        publish the results as required (though possibly without the graphs
        and detailed statistics); \parsername{} provides the database and
        leaves it to the user to utilise it in any way, whenever they want,
        with no restrictions on usage of the data.  The interface is
        optimised for interactive rather than automated queries and it does
        not appear to be possible to write independent tools to utilise the
        Splunk database.  Many types of reports are available, though most
        are variations of a bar or pie chart, with the exception of bubble
        and heatmap graphs.  It is very easy to drill down through the
        graphs to extract a portion of the data (e.g.\ select the hour with
        the largest number of events, then select a particular host, and
        finally a specific address), though it is not possible to search on
        partial words.  The searches are quite fast, though as the free
        version has a limit on the amount of data indexed per day the
        volume of data to be searched was merely 45MB\@; the cheapest
        licensed version costs \$6000, and still limits the volume of data
        indexed per day.

        In the specific case of parsing Postfix logs, Splunk extracts some
        standard fields: to and from addresses, HELO hostname, date, host
        the logs were collected from, and protocol.  It parses the standard
        syslog fields at the beginning of the line, and extracts any
        \texttt{key=<value>} pairs occurring after the standard syslog
        prologue; these pairs are the fields listed above.  Searches can be
        based on the extracted fields, and all text in the line is also
        available for searching.  \parsername{} extracts noticeably more
        data, though it does not make the full text of the line available,
        and the full power of \SQL{} is available when searching, allowing
        the user to search on arbitrarily complicated conditions.

        Splunk is a generic tool, so it lacks any Postfix specific support
        over and above extracting the \texttt{key=<value>} fields; most
        importantly it makes no attempt to correlate log lines by queueid
        or \pid{}, nor to handle any of the myriad complications discussed
        in this document.

@@

        Isoqlog produces a report listing the number of mails sent by each
        unique sender address, and separately the total number of bytes
        transferred; both reports are produced for daily, monthly and
        annual time spans, but only for the domains listed in its
        configuration file (making it impossible to produce reports for
        every sender domain).  It appears to ignore all log lines except
        for those for the current day, though it does maintain a record of
        data previously extracted, which the newly extracted data is merged
        into (no information is provided on the format of the data store).
        The data extracted appears to be limited to the number of mails
        sent by each sender, unlike the copious amounts of data extracted
        by \parsername{}.  It doesn't utilise rejection log lines in any
        way, so is unsuitable for the purposes of this project.  Its
        parsing is completely inextensible, indeed is almost
        incomprehensible, relying on \texttt{scanf(3)}, fixed offsets and
        low level string manipulation; it is the opposite end of the
        spectrum to \parsernames{} parsing.  It doesn't handle any of the
        complications discussed in this document, doesn't gather the
        breadth of data required for this project, and ignores the majority
        of log lines produced by Postfix.

@@

        AWStats will produce simple graphs for many different services, but
        supporting many different services without special purpose code
        restricts it to supporting the \LCD{}.  The data it will extract
        from an \MTA{} log file is limited in comparison to \parsername{}:
        \newline \tab{} time2, email, email\_r, host, host\_r, method, url,
        code and bytesd.\newline  There does not seem to be an explanation
        of any of those fields in the documentation (\parsername{} provides
        copious documentation).  AWStats has no special purpose code to
        deal with the intricacies of Postfix logs, in fact it operates by
        coercing Postfix logs into Apache\footnote{The Apache web server is
        the most popular HTTP server in use for the past 10 years; more
        information is available at \url{http://httpd.apache.org/}.} format
        log files, for analysis by AWStats' HTTP log file parser.  The
        converting parser only deals with a small portion of the log lines
        generated by Postfix, silently skipping those it cannot deal with,
        and does not distinguish between different types of rejection.
        Although though it does correlate log lines by queueid, it does not
        deal with any of the other complications described in this
        document.

@@

\glossary{name={syslog},description={Syslog is the standard logging
    mechanism on Unix systems: the program sends log messages to syslog,
    then syslog filters and stores the messages according to the
    configuration the administrator has chosen.
}}
